{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-06-17 sftrends_manu.Rnw
figures for C10 din

%acronyms
\acrodef{chla}[chl-\textit{a}]{chlorophyll \textit{a}}
\acrodef{sfe}[SFE]{San Francisco Estuary}
\acrodef{wrtds}[WRTDS]{Weighted Regressions on Time, Discharge, and Season}

@@

\subsection{Case studies}
\subsubsection{Disaggregating observed nitrogen time series}
Hypothesis: Because multiple factors influence nutrient concentrations at different times, relationships between nutrients, time, and flow/salinity are non-linear, so we expect 1) annual trend independent of seasonal trend, 2) changes in seasonal amplitudes and quantile trends over time, 3) varying flow contribution, either as difference between predicted/flow-normalized results or changes in nutrient v flow scatterplots at different annual periods.
\subsubsection{Effects of wastewater treatment}
Hypothesis: Modal response of nutrient concentrations at P8 over time is result of WWTP upgrades, so we expect 1) a shift in load contributions before/after upgrade, 2) a flow-normalized annual trend at P8 to show a change concurrent with WWTP upgrades, and 3) shift in the flow/nutrient relatinship before and after upgrade related to change in load contributions.
\subsubsection{Effects of biological invasions}
Hypothesis: Biological invasions by benthic filter feeders have shifted abundance and composition of phytoplankton communities in Suisun Bay, so we expect 1) decline in annual, flow-normalized chlorophyll concentrations over time coincident with increase in abundance of invaders, 2) changes in stoichiometric ratios of limiting nutrients (nitrogen, SiO2) suggesting different uptake rates with shift in community composition, and 3) seasonal shifts in limiting nutrients based on changes in community composition and relative abundances with seasonal succession.

@@

\subsubsection{Clam invasion in Suisun Bay}
Data from \citep{Crauder16}, \citet{Jassby08} describes phytoplankton community changes in the upper estuary, including chlorophyll response to flow.  Figure 10 in \citet{Jassby08} showed that chlorophyll generally decreased with flow in 1980 but inreased with flow in 2000.
\cref{fig:clmchl}, \cref{tab:chlsio}

@@

<<clmchl, fig.height = 7, fig.width = 10, out.width = '\\textwidth', fig.cap = 'Trends in clam abundance and \\ac{chla} concentration from 1976 to 2014 at station D7 in Suisun Bay.  Invasion by \\textit{Potamocorbula amurensis} clams in the late 1980s and displacement of \\textit{Corbicula fluminea} was shown by changes in clam density (top left, annual means).  A coincident decrease in \\ac{chla} concentration was also observed (bottom left).  A weak but significant ($p < 0.001$) relationship between clam biomass and \\ac{chla} concentration is shown in the top right.  Flow relationships with \\ac{chla} concentration have also changed over time (bottom right, observations from June). Chlorophyll shows a slight positive then dominantly negative association with increasing flow (decreasing salinity) early in the time series, whereas the trend is reversed in recent years.'>>=

@@

# annual clam time series
p1 <- ggplot(clmdat, aes(x = yr, y = clams_smp, fill = species)) + 
  geom_bar(stat = 'identity') + 
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    # axis.title.x = element_blank(),
    legend.position = 'top', 
    legend.title = element_blank()
    ) + 
  scale_fill_brewer(
    palette = 'Set2',
    labels = parse(text = c('italic(Corbicula)', 'italic(Potamocorbula)'))
    ) + 
  scale_x_date('', limits = xlims) +
  scale_y_continuous('Clams per sample')

# annual chlorophyll time series with flow-norm
p2 <- ggplot(chldat, aes(x = yr, y = norm0.5)) + 
  geom_point(aes(y = res, shape = 'Observed'), size = 3) + 
  geom_line(aes(color = 'Flow-normalized')) + 
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    # axis.title.x = element_blank(), 
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal'
    ) + 
  scale_colour_manual(values = c('black', 'black')) +
  scale_x_date('', limits = xlims) +
  scale_y_continuous(expression(paste("Chl-", italic(a), " (", italic(mu), "g ", L^-1, ")")))

@@

# scatter plot of concetration x biomass with model 
p3 <- ggplot(toplo, aes(x = biomass, y = norm0.5)) + 
  geom_point(size = 3, alpha = 0.8, colour = 'darkgrey') + 
  geom_line(data = data.frame(x = xvals, y = yvals), aes(x, y, colour = 'black'), lwd = 1) +
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal'
    ) +
  scale_colour_manual(label = 'log(1 + chl) = 1.1 - 0.09log(1 + biomass)', values = 'black') +
  scale_x_continuous(xlab) + 
  scale_y_continuous(ylab, limits = c(0.8, 4))

@@

p4 <- dynaplot(chld7, mo = 6, col_vec = cols, floscl = F) +
  theme_minimal() +
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal',
    strip.background = element_blank(),
    strip.text.x = element_blank()
    ) + 
  scale_x_continuous('ln-Salinity (psu)') + 
  scale_y_continuous(ylab, limits = c(0.4, 3.8))

>>>>

%acronyms
\acrodef{chla}[chl-\textit{a}]{chlorophyll \textit{a}}
\acrodef{din}[DIN]{dissolved inorganic nitrogen}
\acrodef{sfe}[SFE]{San Francisco Estuary}
\acrodef{wrtds}[WRTDS]{Weighted Regressions on Time, Discharge, and Season}

@@

\subsection{Case studies}
These are science questions that are relevant outside of the region.
\subsubsection{Disaggregating observed nitrogen time series}
Hypothesis: Because multiple factors influence nutrient concentrations at different times, relationships between nutrients, time, and flow/salinity are non-linear, so we expect 1) annual trend independent of seasonal trend, 2) changes in seasonal amplitudes and quantile trends over time, 3) varying flow contribution, either as difference between predicted/flow-normalized results or changes in nutrient v flow scatterplots at different annual periods.
\subsubsection{Effects of wastewater treatment}
Hypothesis: Modal response of nutrient concentrations at P8 over time is result of WWTP upgrades, so we expect 1) a shift in load contributions before/after upgrade, 2) a flow-normalized annual trend at P8 to show a change concurrent with WWTP upgrades, and 3) shift in the flow/nutrient relatinship before and after upgrade related to change in load contributions.
\subsubsection{Effects of biological invasions}
Hypothesis: Biological invasions by benthic filter feeders have shifted abundance and composition of phytoplankton communities in Suisun Bay, so we expect 1) decline in annual, flow-normalized chlorophyll concentrations over time coincident with increase in abundance of invaders, 2) changes in stoichiometric ratios of limiting nutrients (nitrogen, SiO2) suggesting different uptake rates with shift in community composition, and 3) seasonal shifts in limiting nutrients based on changes in community composition and relative abundances with seasonal succession.

@@

\subsubsection{Clam invasion in Suisun Bay}
Data from \citep{Crauder16}, \citet{Jassby08} describes phytoplankton community changes in the upper estuary, including chlorophyll response to flow.  Figure 10 in \citet{Jassby08} showed that chlorophyll generally decreased with flow in 1980 but inreased with flow in 2000.

Note the decrease in Potamocorbula abundance in 2011, 2012.  These are wet years where abundance/biomass of the clams is driven down by lower salinity.  Contrased wtih the annual chlorophyll trends in the same years, the predicted values are above the flow-normalized trend suggesting an increase in chlorophyll with higher flow.  The potential mechanism is therefore a decrease in clam abundance with high flow that releases phytoplankon from filtration pressure.  This also explains the positive association of chlorophyll with flow in recent years (bottom right dynaplot).

Further, chlorophyll trends early in the time series generally show a decrease with high flow with a distinct maximum at moderate flow.  This may suggest stratification events at moderate flow contributed to phytoplankton blooms early in the time series. Water withdrawals later in the time series could have also altered environmental conditions to reduce the frequency occurrence of stratification events.  Look into this more...

What about biomass/density relationships for Potamocorbula?  Although clam density increases throughout the period,  What about initial decrease in chlorophyll prior to clam invasion?  Is this related to water withdrawals (i.e., decrease in stratification events at moderate flow)?

\cref{fig:clmchl}, \cref{tab:chlsio}

@@

<<dinc10, fig.height = 8, fig.width = 7, out.width = '0.75\\textwidth', fig.cap = 'Time series of \\ac{din} and flow at station C10.  Subfigure (a) shows the observed \\ac{din} time series and subfigure (b) shows the annual predictions from \\ac{wrtds} at different conditional quantiles ($\\tau$ = 0.1, 0.5, 0.9).  The points in subfigure (b) are predictions of observed \\ac{din} and the lines are flow-normalized predictions.  Subfigure (c) shows the difference between the model predictions and flow-normalized predictions at the fiftieth conditional quantile.  Subfigure (d) shows the flow time series of the San Joaquin River with a locally-estimated (loess) smooth to emphasize the long-term trend.'>>=
# data
data(dinc10)

# din at c10
toplo <- dinc10

# globals
ylab <- expression(paste('DIN (mg ', L^-1, ')'))
ptheme <- theme_minimal() + 
  theme(
    axis.line.x = element_line(),
    axis.line.y = element_line(), 
    axis.title.x = element_blank()
  )
col_vec <- RColorBrewer::brewer.pal(9, 'Set1')[c(3, 2, 1)]
col_lim <- c(0, 3.8)
xlims <- range(toplo$date)
txt <- data.frame(
  x = min(toplo$date), 
  y =  c(5, 2.4, 0.5, 7), 
  txt = c('a', 'b', 'c', 'd')
)
size <- 7

# observed din
toplo1 <- select(toplo, date, res) %>% 
  na.omit()
p1 <- ggplot(toplo1, aes(x = date, y = exp(res))) + 
  geom_line() + 
  ptheme + 
  scale_y_continuous(ylab) +
  theme(
    axis.text.x = element_blank(),
    plot.margin = grid::unit(c(6, 6, -24, 6), 'pt')
  ) +
  geom_text(data = txt[1, ], aes(x = x, y = y, label = txt), size = size) +
  scale_x_date(limits = xlims)

# din prdnrmplot
p2 <- prdnrmplot(toplo, logspace = F, col_vec = col_vec) + 
  ptheme + 
  scale_y_continuous(ylab) +
  theme(
    legend.position = 'top', 
    axis.text.x = element_blank(),
    plot.margin = grid::unit(c(6, 6, -12, 6), 'pt')
  ) +
  geom_text(data = txt[2, ], aes(x = x, y = y, label = txt, colour = NULL), show.legend = F, size = size) +
  scale_x_date(limits = xlims)

# flo/observed diffs
toplo2 <- prdnrmplot(toplo, plot = F)
toplo2 <- left_join(toplo2$fits, toplo2$nrms, by = c('date', 'taus')) %>% 
  mutate(flodiff = fits_value - nrms_value) %>% 
  filter(taus == 0.5)
toplo3 <- toplo
toplo3$flo <- scales::rescale(toplo3$flo, to = attr(toplo, 'floobs_rng'))
ylab <- expression(paste('Pred - norm DIN (mg ', L^-1, ')'))
p3 <- ggplot(toplo2, aes(x = date, y = flodiff)) + 
  geom_bar(stat = 'identity') + 
  theme_minimal() +
  ptheme + 
  theme(
    axis.text.x = element_blank(),
    plot.margin = grid::unit(c(6, 6, -12, 6), 'pt')
    ) + 
  scale_y_continuous(ylab) +
  scale_x_date(limits = xlims) +
  geom_text(data = txt[3, ], aes(x = x, y = y, label = txt), size = size)

# flo with loess
ylab <- expression(paste('ln-Flow (', m^3, ' ', s^-1, ')'))
p4 <- ggplot(toplo3, aes(x = date, y = flo)) + 
  geom_line() +
  stat_smooth(se = F) + 
  ptheme + 
  scale_y_continuous(ylab) +
  scale_x_date(limits = range(toplo3$date)) +
  geom_text(data = txt[4, ], aes(x = x, y = y, label = txt), size = size)

# align widths of plots in first column, first two
pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p2))
pC <- ggplot_gtable(ggplot_build(p3))
pD <- ggplot_gtable(ggplot_build(p4))
maxWidth = grid::unit.pmax(pA$widths[2:3], pB$widths[2:3], pC$widths[2:3], pD$widths[2:3])

pA$widths[2:3] <- maxWidth
pB$widths[2:3] <- maxWidth
pC$widths[2:3] <- maxWidth
pD$widths[2:3] <- maxWidth

grid.arrange(
  pA, pB, pC, pD,
  heights = c(0.9, 1.3, 1, 1.2),
  ncol = 1
  )
@

<<dinc10dyna, fig.height = 6, fig.width = 9, out.width = '\\textwidth', fig.cap = 'Modelled relationships between \\ac{din}, flow, and time at station C10.  The top figure shows the annual and seasonal variation over the entire time series and the bottom figure shows annual variation for selected months to remove seasonal variation.  Warmer colors indicate higher \\ac{din} concentrations.  The y-axis on the bottom figure is truncated by the fifth and ninety-fifth percentiles of flow within each month.  Model results are for the fiftieth conditional quantile of \\ac{din}.'>>=

# data
data(dinc10)

# din at c10
toplo <- dinc10

# globals
ptheme <- theme_minimal() + 
  theme(
    axis.line.x = element_line(),
    axis.line.y = element_line(), 
    axis.title.x = element_blank()
  )
col_lim <- c(0, 2.9)
txt <- data.frame(
  x = min(toplo$date), 
  y =  c(5, 2.4, 0.5, 6), 
  txt = c('a', 'b', 'c', 'd')
)
size <- 7

# dynaplot, all
col_vec = NULL
attr(toplo, 'reslab') <- expression(paste("DIN (mg ", L^-1, ")"))
ylab <- expression(paste('ln-Flow (', m^3, ' ', s^-1, ')'))
p1 <- gridplot(toplo, month = 'all', logspace = F, tau = 0.5, col_lim = col_lim, floscl = F, col_vec = rev(col_vec)) +
  scale_y_continuous(ylab, expand = c(0, 0)) + 
  ptheme + 
  theme(legend.position = 'top')

# dynaplot selected months
p2 <- gridplot(toplo, logspace = F, tau = 0.5, col_lim = col_lim, month = c(1, 4, 7, 10), ncol = 4, floscl = F) + 
  ptheme + 
  theme(
    strip.background = element_blank(),
    legend.position = 'none', 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    ) + 
  scale_y_continuous(ylab, expand = c(0, 0))

# align widths of plots in first column, first two
pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p2))
maxWidth = grid::unit.pmax(pA$widths[2:3], pB$widths[2:3])

pA$widths[2:3] <- maxWidth
pB$widths[2:3] <- maxWidth

grid.arrange(
  pA, pB,
  heights = c(1.2, 1),
  ncol = 1
  )
@

<<clmchl, fig.height = 7, fig.width = 10, out.width = '\\textwidth', fig.cap = 'Trends in clam abundance and \\ac{chla} concentration from 1976 to 2014 at station D7 in Suisun Bay.  Invasion by \\textit{Potamocorbula amurensis} clams in the late 1980s and displacement of \\textit{Corbicula fluminea} was shown by changes in clam density (a, annual means).  A coincident decrease in \\ac{chla} concentration was also observed (c).  A weak but significant ($p < 0.001$) relationship between clam biomass and \\ac{chla} concentration is shown in subfigure (b).  Flow relationships with \\ac{chla} concentration have also changed over time (d, observations from June). Chlorophyll shows a slight positive then dominantly negative association with increasing flow (decreasing salinity) early in the time series, whereas the trend is reversed in recent years.'>>=

@@

# subfig labs
txt <- data.frame(
  x = c(as.numeric(xlims[1]), as.numeric(xlims[2]), 25, 0.25),
  y =  c(10000, 11, 4, 3.8), 
  txt = c('a', 'c', 'b', 'd')
)
size <- 7

# annual clam time series
p1 <- ggplot(clmdat, aes(x = yr, y = clams_smp, fill = species)) + 
  geom_bar(stat = 'identity') + 
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    # axis.title.x = element_blank(),
    legend.position = 'top', 
    legend.title = element_blank()
    ) + 
  scale_fill_brewer(
    palette = 'Set2',
    labels = parse(text = c('italic(Corbicula)', 'italic(Potamocorbula)'))
    ) + 
  scale_x_date('', limits = xlims) +
  scale_y_continuous('Clams per sample') +
  with(txt[1, ], annotate('text', x = xlims[1], y = y, label = txt, size = size))

# annual chlorophyll time series with flow-norm
p2 <- ggplot(chldat, aes(x = yr, y = norm0.5)) + 
  geom_point(aes(y = res, shape = 'Observed'), size = 3) + 
  geom_line(aes(color = 'Flow-normalized')) + 
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    # axis.title.x = element_blank(), 
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal'
    ) + 
  scale_colour_manual(values = c('black', 'black')) +
  scale_x_date('', limits = xlims) +
  scale_y_continuous(expression(paste("Chl-", italic(a), " (", italic(mu), "g ", L^-1, ")"))) +
  with(txt[2, ], annotate('text', x = xlims[2], y = y, label = txt, size = size))

@@

# scatter plot of concetration x biomass with model 
p3 <- ggplot(toplo, aes(x = biomass, y = norm0.5)) + 
  geom_point(size = 3, alpha = 0.8, colour = 'darkgrey') + 
  geom_line(data = data.frame(x = xvals, y = yvals), aes(x, y, colour = 'black'), lwd = 1) +
  theme_minimal() + 
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal'
    ) +
  scale_colour_manual(label = 'log(1 + chl) = 1.1 - 0.09log(1 + biomass)', values = 'black') +
  scale_x_continuous(xlab) + 
  scale_y_continuous(ylab, limits = c(0.8, 4)) +
  geom_text(data = txt[3, ], aes(x = x, y = y, label = txt, colour = NULL), show.legend = F, size = size)

@@

p4 <- dynaplot(chld7, mo = 6, col_vec = cols, floscl = F) +
  theme_minimal() +
  theme(
    axis.line.y = element_line(), 
    axis.line.x = element_line(),
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal',
    strip.background = element_blank(),
    strip.text.x = element_blank()
    ) + 
  scale_x_continuous('ln-Salinity (psu)') + 
  scale_y_continuous(ylab, limits = c(0.4, 3.8)) +
  annotate('text', x = txt[4, 'x'], y = txt[4, 'y'], label = txt[4, 'txt'], size = size)

### 2016-06-17 sftrends_manu.tex
figures for C10 din

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Tue Jun 14 18:01:49 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

The \ac{sfe}... The delta...

\subsection{Analysis method and application}

@@

@@

\caption{Trends in clam abundance and \ac{chla} concentration from 1976 to 2014 at station D7 in Suisun Bay.  Invasion by \textit{Potamocorbula amurensis} clams in the late 1980s and displacement of \textit{Corbicula fluminea} was shown by changes in clam density (top left, annual means).  A coincident decrease in \ac{chla} concentration was also observed (bottom left).  A weak but significant ($p < 0.001$) relationship between clam biomass and \ac{chla} concentration is shown in the top right.  Flow relationships with \ac{chla} concentration have also changed over time (bottom right, observations from June). Chlorophyll shows a slight positive then dominantly negative association with increasing flow (decreasing salinity) early in the time series, whereas the trend is reversed in recent years.}\label{fig:clmchl}
\end{figure}

>>>>

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Thu Jun 16 16:53:30 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

The \ac{sfe}... exchange with ocean, freshwater inputs, runoff variation (drought years), climate, support of natural resources
\subsection{Analysis method and application}

@@

{\centering \includegraphics[width=0.75\textwidth]{figs/dinc10-1}

\caption{Time series of \ac{din} and flow at station C10.  Subfigure (a) shows the observed \ac{din} time series and subfigure (b) shows the annual predictions from \ac{wrtds} at different conditional quantiles ($\tau$ = 0.1, 0.5, 0.9).  The points in subfigure (b) are predictions of observed \ac{din} and the lines are flow-normalized predictions.  Subfigure (c) shows the difference between the model predictions and flow-normalized predictions at the fiftieth conditional quantile.  Subfigure (d) shows the flow time series of the San Joaquin River with a locally-estimated (loess) smooth to emphasize the long-term trend.}\label{fig:dinc10}
\end{figure}

{\centering \includegraphics[width=\textwidth]{figs/dinc10dyna-1}

\caption{Modelled relationships between \ac{din}, flow, and time at station C10.  The top figure shows the annual and seasonal variation over the entire time series and the bottom figure shows annual variation for selected months to remove seasonal variation.  Warmer colors indicate higher \ac{din} concentrations.  The y-axis on the bottom figure is truncated by the fifth and ninety-fifth percentiles of flow within each month.  Model results are for the fiftieth conditional quantile of \ac{din}.}\label{fig:dinc10dyna}
\end{figure}

@@

\caption{Trends in clam abundance and \ac{chla} concentration from 1976 to 2014 at station D7 in Suisun Bay.  Invasion by \textit{Potamocorbula amurensis} clams in the late 1980s and displacement of \textit{Corbicula fluminea} was shown by changes in clam density (a, annual means).  A coincident decrease in \ac{chla} concentration was also observed (c).  A weak but significant ($p < 0.001$) relationship between clam biomass and \ac{chla} concentration is shown in subfigure (b).  Flow relationships with \ac{chla} concentration have also changed over time (d, observations from June). Chlorophyll shows a slight positive then dominantly negative association with increasing flow (decreasing salinity) early in the time series, whereas the trend is reversed in recent years.}\label{fig:clmchl}
\end{figure}
